---
title: "Reproducible Research - Peer Assessment 2"
output: 
  html_document: 
    keep_md: yes
---

#Impact of Severe Weather Evetns on Health and Economy in the United States


##Synopsis
In this report we investigate which type of events are most harmful to population health and economic damage.
This report is based on the U.S National Oceanic and Atmospheric Administration's(NOAA) storm database from 1950-2011.
The results are the estimates of fatalitis, injuries, properties damage costs and crop damage costs to decide which types of event are most harmful to health and economy.
From the results we found that tornado is most harmful to population health, while flood and drought are most harmful to economic consequences.


##Data Processing

### Load Data
```{r}
echo= TRUE
cache = TRUE
library(dplyr)
library(ggplot2)
library(gridExtra)

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(fileUrl, "peer2.bz2", method="curl")
storm <- read.csv(bzfile("peer2.bz2"), header=TRUE)
```

### Calculate the estimate of fatalities and injuries
```{r}
total_fatality <- 
        storm %>% 
        select(EVTYPE, FATALITIES) %>% 
        group_by(EVTYPE) %>% 
        summarise(total_cnt=sum(FATALITIES,na.rm=TRUE)) %>%
        arrange(desc(total_cnt)) %>%
        head(10) %>%
        within(EVTYPE <- factor(EVTYPE, levels=EVTYPE))

total_injury <- 
        storm %>%
        select(EVTYPE, INJURIES) %>%
        group_by(EVTYPE) %>%
        summarise(total_cnt=sum(INJURIES, na.rm=TRUE)) %>%
        arrange(desc(total_cnt)) %>%
        head(10) %>%
        within(EVTYPE <- factor(EVTYPE, levels=EVTYPE))
```

### Calculate the estimate of properties and crops damage costs 
```{r}
toTenPower <- function(x) {
        if(is.numeric(x)) {
                x <- as.numeric(x)
        }
        else if(grepl("h", x, ignore.case=TRUE)) {
                x <-2
        }
        else if(grepl("k", x, ignore.case=TRUE)) {
                x <-3
        }
        else if(grepl("m", x, ignore.case=TRUE)) {
                x <-6
        }
        else if(grepl("b", x, ignore.case=TRUE)) {
                x <-9
        }
        else if(x=="" | x==" ") {
                x <-1
        }
        else {
                x <- 0
        }
        x
}
calculateNum <- function(num, exp) {
        pow <- toTenPower(exp)
        if(is.na(pow)){
                num <-0
        }
        else {
                if(is.numeric(num)) {
                        num <- num*10^pow
                }
                else {
                        num <-0
                }
        }
                
        num
}
storm_prop <- storm %>% select(EVTYPE, PROPDMG, PROPDMGEXP)
storm_prop$propDmage <- mapply(calculateNum, storm_prop$PROPDMG, storm_prop$PROPDMGEXP)
storm_prop$propDmage <- storm_prop$propDmage/(10^9)
storm_crop <- storm %>% select(EVTYPE, CROPDMG, CROPDMGEXP)
storm_crop$cropDmage <- mapply(calculateNum, storm_crop$CROPDMG, storm_crop$CROPDMGEXP)
storm_crop$cropDmage <- storm_crop$cropDmage/(10^9)

total_prop <- 
        storm_prop %>%
        select(EVTYPE, propDmage) %>%
        group_by(EVTYPE) %>%
        summarise(total_cnt=sum(propDmage, na.rm=TRUE)) %>%
        arrange(desc(total_cnt)) %>%
        head(10) %>%
        within(EVTYPE<-factor(EVTYPE, levels=EVTYPE))
total_crop <-
        storm_crop %>%
        select(EVTYPE, cropDmage) %>%
        group_by(EVTYPE) %>%
        summarise(total_cnt=sum(cropDmage, na.rm=TRUE)) %>%
        arrange(desc(total_cnt)) %>%
        head(10) %>%
        within(EVTYPE<-factor(EVTYPE, levels=EVTYPE))
```


##Results

### 1. Across the United States, which types of events are most hamful with respect to population health?
```{r}
drawBarplot <- function(d, title_text, x_text, y_text) {
        ggplot(data=d, aes(x=EVTYPE, y=total_cnt)) + 
                geom_bar(stat="identity") +
                theme(axis.text.x=element_text(angle=30, hjust=1)) +
                labs(title=title_text, x=x_text, y=y_text)
}

figure_fatality <- drawBarplot(total_fatality, "the top 10 highest fatalities", "weather type", "total number")
figure_injury <- drawBarplot(total_injury, "the top 10 hightest injuries", "weather type", "total number")
grid.arrange(figure_fatality, figure_injury, ncol=2)
```

### 2. Across the United States, which types of events have the greatest economic consequences?
```{r}
figure_prop <- drawBarplot(total_prop, "the top 10 highest propertities", "weather type", "cost($ billions)")
figure_crop <- drawBarplot(total_crop, "the top 10 higest crop", "weather type", "cost($ billions)")
grid.arrange(figure_prop, figure_crop, ncol=2)
```

